\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{setspace}
\usepackage{hyperref}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue
}

\definecolor{codebg}{RGB}{245,247,250}
\definecolor{codeborder}{RGB}{220,224,230}

\lstdefinestyle{codestyle}{
    basicstyle=\ttfamily\small,
    backgroundcolor=\color{codebg},
    frame=single,
    rulecolor=\color{codeborder},
    breaklines=true,
    columns=fullflexible,
    keepspaces=true,
    showstringspaces=false,
    xleftmargin=8pt,
    xrightmargin=8pt,
    aboveskip=8pt,
    belowskip=8pt
}

\setstretch{1.08}
\setlist[itemize]{leftmargin=1.4em}
\setlist[enumerate]{leftmargin=1.6em}

\title{Behavior Planner System Instruction and Runtime Prompt}
\author{}
\date{}

\begin{document}

\maketitle

\section{Purpose}

This document records the \emph{current} compact LLM behavior-planner interface
used by the CARLA branch. The system instruction is sent once at startup. After
that, each planner query uses the compact prompt input shown below.

\section{System Instruction Summary}

\subsection{Role}

The model is the high-level behavior planner for the ego vehicle. It must read
the compact environment state and choose one of the allowed behaviors:

\begin{lstlisting}[style=codestyle]
LANE_KEEP
LANE_CHANGE_LEFT
LANE_CHANGE_RIGHT
EMERGENCY_BRAKE
\end{lstlisting}

\subsection{Lane-ID Convention}

\begin{itemize}
    \item The prompt uses a compact lane convention relative to the left edge
    of the road.
    \item Lane $-1$ is the left-most driving lane.
    \item Lane $-2$ is the next lane to the right.
    \item Lane IDs decrease as the lane moves rightward.
\end{itemize}

\subsection{Prompt Fields}

The current system instruction defines the following input blocks:

\begin{enumerate}
    \item \texttt{ID:[prompt\_id]}
    \item \texttt{Ego01:[x,y,v,psi,Llane]}
    \item \texttt{ROUTE:[lane\_safe\_l,lane\_safe\_c,lane\_safe\_r]}
    \item \texttt{PREV:[behavior]}
    \item Zero or more surrounding-vehicle lines:
    \texttt{V[id]:[x,y,v,psi,Llane,I]}
\end{enumerate}

\subsection{Route Safety Meaning}

The current route block is

\begin{lstlisting}[style=codestyle]
ROUTE:[lane_safe_l, lane_safe_c, lane_safe_r]
\end{lstlisting}

with binary values:

\begin{itemize}
    \item \texttt{0}: not safe
    \item \texttt{1}: safe
\end{itemize}

The instruction defines safety as follows:

\begin{itemize}
    \item If a vehicle is \textbf{ahead} in that lane, its longitudinal gap
    must be greater than
    \[
    \text{lookahead distance} + \text{forward safety buffer}.
    \]
    \item If a vehicle is \textbf{behind} in that lane, its longitudinal gap
    must be greater than the configured rear safety distance.
\end{itemize}

\subsection{Decision Policy}

The current instruction gives the model the following decision order:

\begin{enumerate}
    \item Use \texttt{EMERGENCY\_BRAKE} only for a sudden close hazard.
    \item If \texttt{lane\_safe\_c = 1}, choose \texttt{LANE\_KEEP}.
    \item Consider a lane change only if \texttt{lane\_safe\_c = 0}.
    \item If the previous decision was a lane change and the current lane is
    still unsafe, prefer continuing the same lane change.
    \item Repeated \texttt{LANE\_CHANGE\_LEFT} or
    \texttt{LANE\_CHANGE\_RIGHT} responses mean continue the same maneuver,
    not request another extra lane change.
\end{enumerate}

\subsection{Required Output}

The model must return only:

\begin{lstlisting}[style=codestyle]
{
  "id": "PROMPT_ID",
  "behavior": "BEHAVIOR_NAME"
}
\end{lstlisting}

\section{Current Prompt Input Format}

\subsection{Template}

\begin{lstlisting}[style=codestyle]
ID:[prompt_id]

Ego01:[x,y,v,psi,Llane]

ROUTE:[lane_safe_l,lane_safe_c,lane_safe_r]

PREV:[behavior]

V[id_1]:[x,y,v,psi,Llane,I]
V[id_2]:[x,y,v,psi,Llane,I]
...
\end{lstlisting}

\subsection{Field Definitions}

\begin{itemize}
    \item \texttt{x, y}: vehicle position in world coordinates
    \item \texttt{v}: longitudinal speed in m/s
    \item \texttt{psi}: heading angle in radians
    \item \texttt{Llane}: compact prompt lane ID
    \item \texttt{I}: predicted intention code
\end{itemize}

The live surrounding-vehicle intention codes are:

\begin{lstlisting}[style=codestyle]
S  = stationary
KC = lane keeping + constant speed
KA = lane keeping + accelerating
KB = lane keeping + braking
LC = lane change left + constant speed
LA = lane change left + accelerating
LB = lane change left + braking
RC = lane change right + constant speed
RA = lane change right + accelerating
RB = lane change right + braking
\end{lstlisting}

\section{Worked Prompt Example}

\begin{lstlisting}[style=codestyle]
ID:[12]

Ego01:[8.65,-3.185,6.548,0.183481,-3]

ROUTE:[1,1,0]

PREV:[LANE_KEEP]

Vobs_static_workzone:[18.4,-7.4,0,-0.0881,-4,S]
Vcav_waypoint_follow2:[-12.035,-0.476,4,-0.102896,-3,KC]
\end{lstlisting}

In this example:

\begin{itemize}
    \item the left adjacent lane is safe,
    \item the current lane is safe,
    \item the right adjacent lane is not safe,
    \item and the previous decision was \texttt{LANE\_KEEP}.
\end{itemize}

\section{Current Runtime Notes}

\begin{itemize}
    \item The system instruction is loaded from
    \texttt{behavior\_planner/system\_instruction.txt}.
    \item The compact prompt is generated by
    \texttt{behavior\_planner/prompt\_builder.py}.
    \item The response parser expects the returned \texttt{id} to match the
    latest prompt id before accepting the decision.
\end{itemize}

\end{document}
